Generative AI for image reconstruction\\ in Intensity Interferometry: a first attempt

### 2025-07-17 Phase_II/draft/GAN.tex
temporary

\section{Generative Adversarial Networks}
\begin{figure*}[hbt]
	\centering
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=\textwidth]{fig/analysis/Plot_learning_rate_disc_loss.pdf}
		\caption{Discriminator loss for three different learning rates. }
		\label{fig:Plot_learning_rate_discloss}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\centering
		\includegraphics[width=\textwidth]{fig/analysis/Plot_learning_rate_gen_total_loss.pdf}
		\caption{Total generator loss for different learning rates (eqn. (\ref{eq:total_gen_loss})).}
		\label{fig:Plot_learning_rate_genloss}
	\end{subfigure}\hfill
	\caption{Generator and discriminator losses for three different learning rates. The left figure shows the total discriminator loss, and the right figure shows the total generator loss. There is no significant difference, but the highest learning rate might be prone to outliers. }
	\label{fig:Plot_learning_rate_loss}
\end{figure*}
Generative Adversarial Networks (GANs) were introduced by \cite{goodfellow2014generative}. The underlying concept is straightforward: it involves two competing networks. The first network, known as the Generator, produces new images based on an input image. These will be referred to as generated images. The second network, the Discriminator, attempts to distinguish between the generated image (predicted image) and the real image (ground truth).

@@

	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_Kernel_size_disc_loss.pdf}
		\caption{Discriminator loss for three different kernel sizes. }
		\label{fig:Plot_kernel_size_discloss}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_Kernel_size_gen_total_loss.pdf}
		\caption{Total generator loss for three different kernel sizes.}
		\label{fig:Plot_kernel_size_genloss}
	\end{subfigure}\hfill

@@

	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_noise_factor_disc_loss.pdf}
		\caption{Discriminator loss for different noise percentages.}
		\label{fig:Plot_noise_discloss}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_noise_factor_gen_total_loss.pdf}
		\caption{Generator losses for different noise percentages.}
		\label{fig:Plot_noise_genloss}
	\end{subfigure}\hfill

@@

	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_Batchsize_disc_loss.pdf}
		\caption{Discriminator loss for different batch sizes.}
		\label{fig:Plot_batchsize_discloss}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_Batchsize_gen_total_loss.pdf}
		\caption{Generator losses for different batch sizes.}
		\label{fig:Plot_batchsize_genloss}
	\end{subfigure}\hfill

@@

	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_Disc_rep_disc_loss.pdf}
		\caption{Discriminator loss for different Discriminator training.}
		\label{fig:Plot_discrep_discloss}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_Disc_rep_gen_total_loss.pdf}
		\caption{Generator losses for different Discriminator training.}
		\label{fig:Plot_discrep_genloss}
	\end{subfigure}\hfill

@@

	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_N_telescopes_disc_loss.pdf}
		\caption{Discriminator loss for different amounts of telescopes.}
		\label{fig:Plot_telescopes_discloss}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\textwidth]{fig/analysis/Plot_N_telescopes_gen_total_loss.pdf}
		\caption{Generator loss for different amount of telescopes.}
		\label{fig:Plot_telescopes_genloss}
	\end{subfigure}\hfill

>>>>

\section{Generative Adversarial Networks}
\begin{figure*}[hbt]
	%\centering
	%\begin{subfigure}{0.5\linewidth}
	\includegraphics[width=.49\textwidth]{fig/analysis/Plot_learning_rate_disc_loss.pdf}\hfil
	%\caption{Discriminator loss for three different learning rates. }
	%	\label{fig:Plot_learning_rate_discloss}
	%\end{subfigure}\hfill
	%\begin{subfigure}{0.5\linewidth}
	\includegraphics[width=.49\textwidth]{fig/analysis/Plot_learning_rate_gen_total_loss.pdf}
		\caption{Total generator loss for different learning rates (eqn. (\ref{eq:total_gen_loss})).}
	%	\label{fig:Plot_learning_rate_genloss}
	%\end{subfigure}\hfill
	\caption{Generator and discriminator losses for three different learning rates. The left figure shows the total discriminator loss, and the right figure shows the total generator loss. There is no significant difference, but the highest learning rate might be prone to outliers. }
	\label{fig:Plot_learning_rate_loss}
\end{figure*}
Generative Adversarial Networks (GANs) were introduced by \cite{goodfellow2014generative}. The underlying concept is straightforward: it involves two competing networks. The first network, known as the Generator, produces new images based on an input image. These will be referred to as generated images. The second network, the Discriminator, attempts to distinguish between the generated image (predicted image) and the real image (ground truth).

@@

  %\begin{subfigure}{0.5\linewidth}
  \includegraphics[width=.49\textwidth]{fig/analysis/Plot_Kernel_size_disc_loss.pdf}
  %\caption{Discriminator loss for three different kernel sizes. }
  %\label{fig:Plot_kernel_size_discloss}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.5\linewidth}
  \includegraphics[width=.49\textwidth]{fig/analysis/Plot_Kernel_size_gen_total_loss.pdf}
  %\caption{Total generator loss for three different kernel sizes.}
  %\label{fig:Plot_kernel_size_genloss}
  %\end{subfigure}\hfill

@@

  %\begin{subfigure}{0.5\linewidth}
  \includegraphics[width=.49\textwidth]{fig/analysis/Plot_noise_factor_disc_loss.pdf}
  %\caption{Discriminator loss for different noise percentages.}
  %\label{fig:Plot_noise_discloss}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.5\linewidth}
  %\centering
  \includegraphics[width=.49\textwidth]{fig/analysis/Plot_noise_factor_gen_total_loss.pdf}
  %\caption{Generator losses for different noise percentages.}
  %\label{fig:Plot_noise_genloss}
  %\end{subfigure}\hfill

@@

  %\begin{subfigure}{0.5\linewidth}
  \includegraphics[width=.49\textwidth]{fig/analysis/Plot_Batchsize_disc_loss.pdf}\hfill
  %\caption{Discriminator loss for different batch sizes.}
  %\label{fig:Plot_batchsize_discloss}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.5\linewidth}
  \includegraphics[width=.49\textwidth]{fig/analysis/Plot_Batchsize_gen_total_loss.pdf}
  %\caption{Generator losses for different batch sizes.}
  %\label{fig:Plot_batchsize_genloss}
  %\end{subfigure}\hfill

@@

  %\begin{subfigure}{0.5\linewidth}
  \includegraphics[width=.49\textwidth]{fig/analysis/Plot_Disc_rep_disc_loss.pdf}\hfil
  %\caption{Discriminator loss for different Discriminator training.}
  %\label{fig:Plot_discrep_discloss}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.5\linewidth}
  \includegraphics[width=.49\textwidth]{fig/analysis/Plot_Disc_rep_gen_total_loss.pdf}
  %\caption{Generator losses for different Discriminator training.}
  %\label{fig:Plot_discrep_genloss}
  %\end{subfigure}\hfill

@@

  %\begin{subfigure}{0.5\linewidth}
  \includegraphics[width=.49\textwidth]{fig/analysis/Plot_N_telescopes_disc_loss.pdf}\hfil
  %\caption{Discriminator loss for different amounts of telescopes.}
  %\label{fig:Plot_telescopes_discloss}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.5\linewidth}
  \includegraphics[width=\textwidth]{fig/analysis/Plot_N_telescopes_gen_total_loss.pdf}\hfil
  %\caption{Generator loss for different amount of telescopes.}
  %\label{fig:Plot_telescopes_genloss}
  %\end{subfigure}\hfill

### 2025-07-17 Phase_II/draft/II.tex
temporary

	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft.jpg}
		\caption{The signal (absolute value of the Fourier transform of the source).}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft_log.jpg}
		\caption{The logarithm of the signal.}
	\end{subfigure}
	\caption{Absolute value of the two-dimensional Fast Fourier Transform of Fig.~\ref{fig:image} measured by Intensity Interferometry, and observation of maximum $(u, v)$ plane with the assumed telescope separations can be completed with the help of Earth rotation.}

@@

	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft_base.jpg}
		\caption{The signal as samples by the baselines.}
	\end{subfigure}\hfill
	\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=\linewidth]{fig/ft/ft_log_base.jpg}
		\caption{The logarithmic signal as sampled by the baselines.}
	\end{subfigure}

>>>>

	%\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=.49\linewidth]{fig/ft/ft.jpg}\hfil
	%\caption{The signal (absolute value of the Fourier transform of the source).}
	%\end{subfigure}\hfill
	%\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=.49\linewidth]{fig/ft/ft_log.jpg}
		%\caption{The logarithm of the signal.}
	%\end{subfigure}
	\caption{Absolute value of the two-dimensional Fast Fourier Transform of Fig.~\ref{fig:image} measured by Intensity Interferometry in the $(u, v)$ plane. \textit{Left.}~Signal on a linear scale. \text{Right}.~Signal on a logarithmic scale.}

@@

	%\centering
	%\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=.49\linewidth]{fig/ft/ft_base.jpg}\hfil
		%\caption{The signal as samples by the baselines.}
	%\end{subfigure}\hfill
	%\begin{subfigure}{0.5\linewidth}
		\includegraphics[width=.49\linewidth]{fig/ft/ft_log_base.jpg}
		%\caption{The logarithmic signal as sampled by the baselines.}
	%\end{subfigure}

### 2025-07-17 Phase_II/draft/Result.tex
rearrange
temporary

\subsection{Predicted Image from the Trained GAN}
Fig.~\ref{fig:GAN} demonstrates the success of the GAN in training a model for Intensity Interferometry (II) to reconstruct images of fast-rotating stars. The GAN was trained on the training datasets for 60,000 steps and subsequently tested on various validation datasets to produce predicted images of a fast-rotating star. In Fig.~\ref{fig:GAN}, four combined images illustrate the GAN's performance in reconstructing the star’s shape, size, and brightness distribution using II.
\begin{itemize}
\item{The left panel shows the signals collected from six baselines, which serve as the input for the Generator during training.}
\item{The first middle panel displays the real image, or ground truth, which the Discriminator uses to distinguish between real images and those generated by the Generator. During training, the GAN aims to replicate these ground truth images.}
\item{The second middle panel presents the reconstructed, or predicted, image produced by the trained GAN, highlighting its success in image reconstruction.}
\item{The right panel shows the difference between the ground truth and the predicted image, with a smaller difference indicating higher precision in image reconstruction.}
\end{itemize}
The predicted images in Fig.~\ref{fig:GAN} yield encouraging results, accurately conveying visual information about the source's size, shape, and brightness distribution across its surface using only six baselines. However, further improvements can be achieved by increasing the number of telescopes to maximize coverage of the $(u, v)$ plane, making the existing and upcoming CTAO an ideal candidate for this approach.
\subsection{Evaluation of GAN using Moments}
The reconstructed images are visually compelling, demonstrating the GAN's effectiveness in using II to reconstruct images. However, visual assessment alone is insufficient; statistical evaluation is necessary to validate the results. To achieve this, we employ image moments as a statistical method. Image moments capture key properties of the reconstructed objects—such as shape, size, and intensity distribution—by quantifying features like position, orientation, and brightness distribution. By comparing the moments of the GAN-generated images to those of the ground truth, we can objectively assess the consistency and accuracy of the reconstruction. This approach provides a reliable framework for evaluating reconstruction quality, as image moments can reveal subtle differences in geometric and intensity properties that might not be apparent through visual inspection alone.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Furthermore, these calculated centroids are instrumental in analyzing the shape, size, and brightness distribution of fast-rotating stars using higher-order image moments. To this end, the central moment of an image is calculated according to:
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - x_c)^p (y - y_c)^q I(x, y).
\end{equation}
The sum of \(p\) and \(q\) defines the order of the central moment.

>>>>

\subsection{Predicted Image from the Trained GAN}
Fig.~\ref{fig:GAN} demonstrates the success of the GAN in training a model for Intensity Interferometry (II) to reconstruct images of fast-rotating stars. The GAN was trained on the training datasets for 60,000 steps and subsequently tested on various validation datasets to produce predicted images of a fast-rotating star. In Fig.~\ref{fig:GAN}, four combined images illustrate the GAN's performance in reconstructing the star’s shape, size, and brightness distribution using II.
\begin{itemize}
\item{The left panel shows the signals collected from six baselines, which serve as the input for the Generator during training.}
\item{The first middle panel displays the real image, or ground truth, which the Discriminator uses to distinguish between real images and those generated by the Generator. During training, the GAN aims to replicate these ground truth images.}
\item{The second middle panel presents the reconstructed, or predicted, image produced by the trained GAN, highlighting its success in image reconstruction.}
\item{The right panel shows the difference between the ground truth and the predicted image, with a smaller difference indicating higher precision in image reconstruction.}
\end{itemize}
The predicted images in Fig.~\ref{fig:GAN} yield encouraging results, accurately conveying visual information about the source's size, shape, and brightness distribution across its surface using only six baselines. However, further improvements can be achieved by increasing the number of telescopes to maximize coverage of the $(u, v)$ plane, making the existing and upcoming CTAO an ideal candidate for this approach.

\subsection{Evaluation of GAN using Moments}
The reconstructed images are visually compelling, demonstrating the GAN's effectiveness in using II to reconstruct images. However, visual assessment alone is insufficient; statistical evaluation is necessary to validate the results. To achieve this, we employ image moments as a statistical method. Image moments capture key properties of the reconstructed objects—such as shape, size, and intensity distribution—by quantifying features like position, orientation, and brightness distribution. By comparing the moments of the GAN-generated images to those of the ground truth, we can objectively assess the consistency and accuracy of the reconstruction. This approach provides a reliable framework for evaluating reconstruction quality, as image moments can reveal subtle differences in geometric and intensity properties that might not be apparent through visual inspection alone.

@@

\begin{figure*}
  %\begin{subfigure}{0.33\linewidth}
  \includegraphics[width=.32\linewidth]{fig/moments/mom3.png}\hfil
  %\caption{The 2nd order moment $\mu_{11}$.}
  %\label{fig:mom4}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.33\linewidth}
  \includegraphics[width=.32\linewidth]{fig/moments/mom4.png}\hfil
  %\caption{The 2nd order moment $\mu_{20}$.}
  %\label{fig:mom5}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.33\linewidth}
  \includegraphics[width=.32\linewidth]{fig/moments/mom5.png}
  %\caption{The 2nd order moment $\mu_{02}$.}
  %\label{fig:mom6}
  %\end{subfigure}\hfill
  \caption{The second-order central moments provide information about the size and shape of stellar objects.  Shown here are all the second-order central moments for ground truth and predicted images generated by the trained GAN. From left to right these are $\mu_{11},\mu_{20},\mu_{02}$.}
	\label{fig:struc}
\end{figure*}
\begin{figure*}
  %\begin{subfigure}{0.50\linewidth}
  \includegraphics[width=.49\linewidth]{fig/moments/mom6.png}\hfil
  %\caption{The 3rd order moment $\mu_{30}$.}
  %\label{fig:mom7}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.50\linewidth}
  \includegraphics[width=.49\linewidth]{fig/moments/mom7.png}
  %\caption{The 3rd order moment $\mu_{03}$.}
  %\label{fig:mom8}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.50\linewidth}
  \includegraphics[width=.49\linewidth]{fig/moments/mom8.png}
  %\caption{The 3rd order moment $\mu_{21}$.}
  %\label{fig:mom9}
  %\end{subfigure}\hfill
  %\begin{subfigure}{0.50\linewidth}
  \includegraphics[width=.49\linewidth]{fig/moments/mom9.png}
  %\caption{The 3rd order moment $\mu_{12}$.}
  %\label{fig:mom10}
  %\end{subfigure}\hfill
  \caption{Shown here are all the third-order central moments for
    ground truth and predicted images generated by the trained GAN.
    They represent the skewness of the brightness distributions.  The
    panels in reading order show
    $\mu_{30},\mu_{03},\mu_{21},\mu_{12}$.}
  \label{fig:moments}
\end{figure*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Furthermore, these calculated centroids are instrumental in analyzing the shape, size, and brightness distribution of fast-rotating stars using higher-order image moments. To this end, the central moment of an image is calculated according to:
\begin{equation}
	\mu_{pq} = \frac{1}{M_{00}}\sum_{x} \sum_{y} (x - x_c)^p (y - y_c)^q I(x, y).
\end{equation}
The sum of \(p\) and \(q\) defines the order of the central moment.

### 2025-07-17 Phase_II/draft/main.tex
rearrange
temporary

\documentclass[fleqn,usenatbib,twocolumn]{mnras}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{mathtools}
\usepackage{siunitx}
\usepackage{enumerate}
\usepackage{caption,subcaption}
\usepackage{comment}
\captionsetup{compatibility=false}
\usepackage{animate}
\usepackage{amsmath} % For \text and many other math symbols
\usepackage{amssymb} % For \mathbb
\usepackage{amsfonts} % For \mathbb
%%%%%
\usepackage{color}
%\usepackage[normalem]{ulem}
\usepackage{soul}
\usepackage{ulem}
\usepackage{comment}

\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\norm}[2]{\left\lVert#1\right\rVert}

\title{Generative AI for image reconstruction in Intensity Interferometry: a first attempt}
\date{}

\begin{document}
\maketitle

\begin{abstract}
In the last few years Intensity Interferometry has made significant strides in achieving high-precision resolution of stellar objects at optical wavelengths. Despite these advancements, phase retrieval remains a major challenge due to the nature of photon correlation. This paper explores the application of a conditional Generative Adversarial Network (cGAN) to tackle the problem of image reconstruction in Intensity Interferometry. This approach successfully reconstructs the shape, size, and brightness distribution of a fast-rotating star from sparsely sampled, spatial power spectrum of the source, corresponding to II with four telescopes.  Although this particular example could also be addressed using parameter fitting, the results suggest that with larger arrays much more complicated systems could be reconstructed by applying machine-learning techniques to II.
\end{abstract}
\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}

\bibliographystyle{mnras}
\bibliography{refs}

>>>>

\documentclass[twocolumn]{emulateapj}

\usepackage{amsmath}

\newcommand{\mias}{$\SI{}{\micro{\rm as}}$}
\newcommand{\abs}[1]{\left|#1\right|}
\newcommand{\norm}[2]{\left\lVert#1\right\rVert}

\begin{document}

\title{Generative AI for image reconstruction\\ in Intensity Interferometry: a first attempt}

\author{Km Nitu Rai,$^{1,2}$, Yuri van der Burg$^{3}$, Soumen
  Basak$^{1}$, Prasenjit Saha,$^{3}$ and Subrata Sarangi$^{4,5}$}

\affiliation{$^{1}$School of Physics, Indian Institute of Science
  Education and Research Thiruvananthapuram, Maruthamala PO, Vithura,
  Thiruvananthapuram 695551, Kerala, India.}

\affiliation{$^{2}$Aryabhatta Research Institute of Observational Sciences, Manora Peak, Nainital 263129, India.}

\affiliation{$^{3}$Physik-Institut, University of Zurich,
  Winterthurerstrasse 190, 8057 Zurich, Switzerland.}

\affiliation{$^{4}$School of Applied Sciences, Centurion University of
  Technology and Management, Odisha-752050, India.}

\affiliation{$^5$ Visiting Associate, Inter-University Centre for Astronomy and Astrophysics, Post Bag 4, Ganeshkhind, Pune 411 007, Maharashtra, India.}

\begin{abstract}
In the last few years Intensity Interferometry has made significant strides in achieving high-precision resolution of stellar objects at optical wavelengths. Despite these advancements, phase retrieval remains a major challenge due to the nature of photon correlation. This paper explores the application of a conditional Generative Adversarial Network (cGAN) to tackle the problem of image reconstruction in Intensity Interferometry. This approach successfully reconstructs the shape, size, and brightness distribution of a fast-rotating star from sparsely sampled, spatial power spectrum of the source, corresponding to II with four telescopes.  Although this particular example could also be addressed using parameter fitting, the results suggest that with larger arrays much more complicated systems could be reconstructed by applying machine-learning techniques to II.
\end{abstract}

\maketitle

\input{intro}
\input{II}
\input{GAN}
\input{ParaGan}
\input{Result}
\input{conclusion}

\bibliographystyle{aa}
\bibliography{refs}
